%-------------------------
% Resume in Latex
% Author : Jake Gutierrez
% Based off of: https://github.com/sb2nov/resume
% License : MIT
%------------------------

\documentclass[letterpaper,11pt]{article}

\usepackage{latexsym}
\usepackage[empty]{fullpage}
\usepackage{titlesec}
\usepackage{marvosym}
\usepackage[usenames,dvipsnames]{color}
\usepackage{verbatim}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
\usepackage{fancyhdr}
\usepackage[english]{babel}
\usepackage{tabularx}
\input{glyphtounicode}
%\sloppy   % reduces overfull hbox warnings on narrow lines
%\usepackage{microtype}  % optional, nicer spacing (pdfLaTeX)
\usepackage{ragged2e}
\usepackage[normalem]{ulem}
%----------FONT OPTIONS----------
% sans-serif
% \usepackage[sfdefault]{FiraSans}
% \usepackage[sfdefault]{roboto}
% \usepackage[sfdefault]{noto-sans}
% \usepackage[default]{sourcesanspro}

% serif
% \usepackage{CormorantGaramond}
% \usepackage{charter}


\pagestyle{fancy}
\fancyhf{} % clear all header and footer fields
\fancyfoot{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

% Adjust margins
\addtolength{\oddsidemargin}{-0.5in}
\addtolength{\evensidemargin}{-0.5in}
\addtolength{\textwidth}{1in}
\addtolength{\topmargin}{-.5in}
\addtolength{\textheight}{1.0in}

\urlstyle{same}

\raggedbottom
\raggedright
\setlength{\tabcolsep}{0in}

% Sections formatting
\titleformat{\section}{
  \vspace{-4pt}\scshape\raggedright\large
}{}{0em}{}[\color{black}\titlerule \vspace{-5pt}]

% Ensure that generate pdf is machine readable/ATS parsable
\pdfgentounicode=1

%-------------------------
% Custom commands
\newcommand{\resumeItem}[1]{
  \item\small{
    {#1 \vspace{-2pt}}
  }
}

\newcommand{\resumeSubheading}[4]{
  \vspace{-2pt}\item
    \begin{tabular*}{0.97\textwidth}[t]{l@{\extracolsep{\fill}}r}
      \textbf{#1} & #2 \\
      \textit{\small#3} & \textit{\small #4} \\
    \end{tabular*}\vspace{-7pt}
}

\newcommand{\resumeSubSubheading}[2]{
    \item
    \begin{tabular*}{0.97\textwidth}{l@{\extracolsep{\fill}}r}
      \textit{\small#1} & \textit{\small #2} \\
    \end{tabular*}\vspace{-7pt}
}

\newcommand{\resumeProjectHeading}[2]{
    \item
    \begin{tabular*}{0.97\textwidth}{l@{\extracolsep{\fill}}r}
      \small#1 & #2 \\
    \end{tabular*}\vspace{-7pt}
}

\newcommand{\resumeSubItem}[1]{\resumeItem{#1}\vspace{-4pt}}

\renewcommand\labelitemii{$\vcenter{\hbox{\tiny$\bullet$}}$}

\newcommand{\resumeSubHeadingListStart}{\begin{itemize}[leftmargin=0.15in, label={}]}
\newcommand{\resumeSubHeadingListEnd}{\end{itemize}}
\newcommand{\resumeItemListStart}{\begin{itemize}}
\newcommand{\resumeItemListEnd}{\end{itemize}\vspace{-5pt}}

%-------------------------------------------
%%%%%%  RESUME STARTS HERE  %%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{document}

\begin{center}
    \textbf{\LARGE \scshape Varshini Prakash}
\end{center}

\begin{center}
    

\href{mailto:varshini.prakash@gmail.com}{\uline{varshini.prakash@gmail.com}} $|$
\href{https://varshini-prakash.github.io}{\uline{varshini-prakash.github.io}} $|$ \href{https://www.linkedin.com/in/varshini-prakash/}{\uline{LinkedIn}} 
\end{center}

\section{Summary}
%Machine Learning Engineer with a strong foundation in deep learning and transformer models. Experienced in building ML pipelines and deploying production-ready models using Python, PyTorch, and TensorFlow.
\begin{justify}
Machine Learning Engineer with a strong foundation in deep learning, transformer models, and applied AI systems. Experienced in building ML pipelines, experimentation workflows, and deploying production-ready models using Python, PyTorch, and TensorFlow.
\end{justify}

%-----------EDUCATION-----------
\section{Education}
  \resumeSubHeadingListStart
    \resumeSubheading
      {MSc Computing Science}{University of Alberta, Canada}
      {GPA 3.9, Thesis Advisor: Osmar Zaiane}{2024}
      \resumeSubheading
      {Bachelor of Technology in Computer Science}{VIT University, Vellore, India}{GPA: 9.26 / 10.0}{2020}
  \resumeSubHeadingListEnd

%-----------Experience-----------
\section{Experience}
\begin{justify}
  \resumeSubHeadingListStart

    \resumeSubheading
      {Mood Disorders Society of Canada}{2024 -- 2026}
      {Machine Learning Engineer}{Edmonton, Canada}

\resumeItemListStart

        \resumeItem{Designed and deployed \textbf{MIRA}, a mental health assistant combining Retrieval-Augmented Generation \textbf{(RAG)} with transformer embeddings and scalable inference to deliver context-aware resource recommendations.}
        \resumeItem{Shipped \textbf{layered safety guardrails}: toxic-content detection, structured refusal paths, fallbacks, and conversational constraints for sensitive-domain use.}
        \resumeItem{Built \textbf{data ingestion and serving workflows} for retrieval, embedding generation, and production inference; standardized release practices for model and index updates.}
        \resumeItem{Ran an \textbf{offline evaluation harness} using curated query sets and systematic sweeps over similarity thresholds, retrieval parameters, and prompt templates, to iteratively refine response quality.}
        \resumeItem{Maintained legacy \textbf{Rasa}-based MIRA; integrated \textbf{DIET} (Dual Intent and Entity Transformer) to strengthen intent recognition and entity extraction in rule-driven flows.}
        \resumeItem{Extended the domain \textbf{knowledge graph} with representation learning and link prediction, improving coverage for semantic retrieval and downstream recommendations.}

\resumeItemListEnd
 
        \resumeSubheading
      {University of Alberta}{2021 -- 2024}
      %{Sept. 2021 -- Present}
      {Graduate Research Assistant Fellow}{Edmonton, Canada}
      \resumeItemListStart

\resumeItem{With Dr.\ Osmar Zaiane, built a \textbf{no-code chatbot builder} enabling configurable dialogue workflows; implemented graph-based flow visualization and dynamic routing for scalable, deployable conversational systems.}
\resumeItem{With Dr.\ Alona Fyshe, developed a \textbf{neural decoding pipeline} using transformer embeddings (BERT, RoBERTa) to study metaphor processing in fMRI data; evaluated alignment between brain activation patterns and contextualized semantic embeddings using similarity and correlation metrics.}
\resumeItemListEnd

\resumeSubheading
{Cognitive Systems Lab}{2019}
%{Sept. 2021 -- Present}
{Summer Research Intern}{Bremen, Germany}
\resumeItemListStart
\resumeItem{Collaborated with Prof. Tanja Schultz to train an EMG-Speech system for Silent Speech, to align lip-reading data from images with phonemes of corresponding EMG signals using 3D Convolutional Neural Networks (CNNs).}
\resumeItemListEnd

\resumeSubheading
{One Fourth Labs, IIT Madras}{2018}
{Content Development Intern}
{Chennai, India}
\resumeItemListStart
\resumeItem{Authored hands-on Deep Learning coursework and programming assignments aligned to industry practices, with Dr. Mitesh Khapra and Dr. Pratyush Kumar.}
\resumeItemListEnd

        %{Designed programming assignments for an online course on Deep Learning. Collaborated with Dr Mitesh Khapra and Dr Pratyush Kumar to ensure content aligned with industry-relevant ML practices.}

            \resumeSubheading
        {Siam Computing}{2018}
        {Chatbot Developer}
        {Chennai, India}
        \resumeItemListStart
        \resumeItem{Designed a conversational chatbot using Google Dialogflow for a retail enterprise, enabling context-aware responses and handling multi-turn dialogues to enhance customer interaction efficiency.}

        \resumeItemListEnd

  \resumeSubHeadingListEnd
\end{justify}


%-----------Extracurriculars-----------
\iffalse
\section{Publications and Posters}
  \resumeSubHeadingListStart

    \resumeSubheading
    {{\bf Prakash, V.}, Chouinard, B., \& Fyshe, A. (2023). Decoding Neural Representations of Sentences in Individuals with Autism. Poster presented at the Reverse Expo, University of Alberta.}
    \resumeSubheading
    {{\bf Prakash, V.}, Moorthy, K., \& Jose, J. T., (2020). Enhancing Predictability of Handwritten Document Content using HTR and Word Substitution. IJISME. }
    \resumeSubheading
    {{\bf Prakash, V.}, \& Tripathy, B. K. (2020). Recent Advancements in Automatic Sign Language Recognition (SLR). In {\em Computational Intelligence for Human Action Recognition} (pp. 1-24). Chapman and Hall/CRC. }


  \resumeSubHeadingListEnd
  \fi



%-----------Publications-----------
\section{Publications and Posters}
\begin{justify}
\resumeItemListStart

\resumeItem{
{\bf Prakash, V.}, Ali Gharaat, M., Lambe Foster, A., Merrick, D., Desnoyers, E., Noble, J. M., \& Zaiane, O. R. (2025).
\textit{Mental Health Resource Retrieval Using Semantic Similarity and Knowledge Graphs.}
International Conference on Digital Economy (pp. 89--98).
{\bf Outstanding Paper Award.}
}
\resumeItem{
{\bf Prakash, V.}, Foster, A. L., Noble, J., \& Zaiane, O. (2024).
\textit{Integrating Conversational Pathways with a Chatbot Builder Platform.}
Information Integration and Web Intelligence: 26th International Conference.
}
\resumeItem{
{\bf Prakash, V.}, Chouinard, B., \& Fyshe, A. (2023).
\textit{Decoding Neural Representations of Sentences in Individuals with Autism.}
Poster presented at the Reverse Expo, University of Alberta.
}
\resumeItem{
{\bf Prakash, V.}, \& Tripathy, B. K. (2020).
\textit{Recent Advancements in Automatic Sign Language Recognition (SLR).}
In \textit{Computational Intelligence for Human Action Recognition} (pp. 1--24).
}

\resumeItemListEnd

\end{justify} 


\iffalse
\noindent
{\bf Prakash, V.}, Moorthy, K., \& Jose, J. T. (2020). Enhancing Predictability of Handwritten Document Content using HTR and Word Substitution. {\em IJISME}. 

\vspace{0.5em}
\fi








\iffalse
%-----------Courses-----------
\section{Classes}
 \begin{itemize}[leftmargin=0.15in, label={}]
    \small{\item{
    \vspace{1mm}
     \textbf{Courses}{: AP Calulus BC, AP Chemistry, AP Physics, AP World History, AP U.S. History, AP Statistics, AP Literature} \\

    }}
 \end{itemize}
 \fi



\section{Awards}
\begin{justify}
\resumeItemListStart
\resumeItem{Outstanding Paper Award at International Conference on Digital Economy (ICDEc) 2025.}
\resumeItem{Certificate of Accomplishment for 16th Annual Student Unmanned Air Systems Competition. }
\resumeItem{Certificate of Appreciation for oral presentation on Data Security for IoT-Based Healthcare System - A cryptographic approach, ICWRS \& ICAFEE 2018.} 
\resumeItem{Recipient of the Special Student Achiever Award from VIT University Vellore (2018-2020).}
\resumeItem{Finalist in HSBC Artificial Intelligence Hackathon conducted by HSBC and IIT Madras.}
\resumeItemListEnd
\end{justify}
%-----------Skills-----------
\section{Skills}
\begin{itemize}[leftmargin=0.15in, label={}]
    \small{\item{
    \textbf{Programming:} Python, C++, SQL \\
    \textbf{ML/DL:} PyTorch, TensorFlow, HuggingFace Transformers, Scikit-learn, NLP \\
    \textbf{LLM Systems:} RAG pipelines, Vector Search, Model Context Protocol (MCP), Prompt Engineering \\
    \textbf{Data \& Scientific Tools:} NumPy, Pandas, Jupyter Notebooks, Conda \\
    \textbf{Systems \& Deployment:} GPU/CUDA environments, Docker, Git, REST APIs, Linux \\
    \textbf{Graph ML:} Knowledge Graphs, GNNs, Representation Learning, Semantic Search
    }}
\end{itemize}

\section{Selected Projects}
\begin{justify}

\iffalse
\resumeSubheading{MIRA Mental Health Chatbot}{2024}{}{} \vspace{-0.75em}
\resumeItemListStart
\resumeItem{\textbf{Safety} — Added layered safety guardrails, including persona-specific prompt constraints, structured refusal logic, and a QwenGuard classifier to detect unsafe or self-harm content before responses reach users.}

\resumeItem{\textbf{Explainability} — Designed and implemented a real-time context feature that dynamically highlights key entities and semantic concepts from conversations, allowing users to trace model reasoning and improving AI transparency.}

\resumeItemListEnd
\fi

\resumeSubheading{Multimodal Speech Recognition System} {2025}{}{} \vspace{-0.75em} 
\resumeItemListStart
\resumeItem{Built an end-to-end audio-visual speech recognition system in PyTorch on GRID using a ResNet-18 lip encoder, mel-spectrogram CNN, gated multimodal fusion (temporal alignment + learned gates), and a CTC-trained BiGRU decoder.}\vspace{-1.35em} \resumeItem{Evaluated WER/CER on speaker-held-out splits with audio/video ablations and noisy-audio stress tests; shipped a reproducible training/eval pipeline with checkpointing, TensorBoard logging, and CTC beam search.}
\resumeItemListEnd

\resumeSubheading{Noise Eliminator for Active Learning}{2021}{}{} \vspace{-0.75em}
\resumeItemListStart \resumeItem{An interactive machine learning-based algorithm that helps produce automated systems that provide highly accurate predictions, with a low number of training instances.}    \resumeItemListEnd

\resumeSubheading{Stock Forecasting}{2021}{}{} \vspace{-0.75em}
\resumeItemListStart \resumeItem{Developed a stock forecasting model framed as a binary classification task to identify profitable stocks. Trained several classifiers on historical data and leveraged their predictions to build portfolios that surpass benchmark performance.}    \resumeItemListEnd


\resumeSubheading{Unmanned Aerial System}{2019}{}{} \vspace{-0.95em}

\resumeItemListStart
\resumeItem{Competed in AUVSI SUAS 2018 (Maryland, USA) as team Ardra; integrated and field-demonstrated a full unmanned aerial system (UAS) for autonomous mission execution.}\vspace{-0.3em}
\resumeItem{Developed interoperability modules and real-time obstacle visualization to support navigation, system coordination, and mission awareness.}
\resumeItemListEnd

\end{justify}

\iffalse
\section{Community Involvement}


\resumeSubheading{Unit Leader}{2024-Present}{Girl Guides of Canada}{} \vspace{0.5em}
\resumeItemListStart {Mentor and lead a group of girls through weekly Girl Guides programs that foster leadership, teamwork, and skill-building. Organize outdoor activities, community service, and badge-based learning experiences.}
\resumeItemListEnd

\resumeSubheading{CSGSA Treasurer}{2021-2022}{Computing Science Graduate Students' Association}{University of Alberta} \vspace{0.5em}
\resumeItemListStart {Oversaw CSGSA finances, including budgeting, expense tracking, and reporting. Collaborated with student leadership to fund social, and professional development events for the graduate student community.}
\resumeItemListEnd



\resumeSubheading{GSA Equity, Diversity, and Inclusion Committee Member}{2021-2022}{Graduate Students' Association}{University of Alberta} \vspace{0.5em}
\resumeItemListStart {Contributed to initiatives that advance Equity, Diversity, and Inclusion (EDI) within the GSA, promoting a welcoming and inclusive community that enhances the graduate student experience on campus.}
\resumeItemListEnd
\fi

\iffalse
  \resumeItemListStart  
    \resumeItem{Built and deployed a production-grade RAG-based mental health assistant using transformer embeddings and vector search, improving resource retrieval accuracy.} 

    \resumeItem{Optimized a handcrafted knowledge graph using deep learning models to predict missing links, refine structure, and generate embeddings to improve RAG.}
    
    \resumeItem{Developed and maintained MIRA, an AI mental health assistant, using the Rasa framework and transformer-based models for intent and entity detection.}  
     
    \resumeItem{Implemented scalable data pipelines and retrieval systems to improve resource recommendation accuracy, using embeddings and vector search.}  

    \resumeItem{Optimized chatbot performance through prompt engineering, similarity threshold tuning, and iterative evaluation, improving response relevance and user engagement metrics.}

    \resumeItem{Implemented and experimented with state-of-the-art safety guardrails, including toxic content detection, fallback mechanisms, and conversational constraints, to ensure responsible AI interactions in sensitive mental health contexts.}

    \resumeItemListEnd 

    \fi

\end{document}